Pre-/Post-Trikafta in CFRD

### 2022-04-28 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
colleen wood data mostly cleaned

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lmerTest)
library(nlme)
library(broom.mixed)
library(segmented)
library(performance)
library(AGD)
library(rspiro)
library(lubridate)
library(plotly)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

@@

@@

@@

@@

```{r}
# Most recent before/after - limit to 2 years before start
# Want the value closest to start date before start, and most recent with respect to now from start
# (i.e. the longest time after start)
most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
  mutate(time_new = max(Time) - Time) %>%
  arrange(time_new) %>% 
  summarise(BMIz = first(na.omit(BMIz)),ppFEV = first(na.omit(ppFEV)),
            ppFVC = first(na.omit(ppFVC))) %>%
  arrange(Site,CFF.ID,Before_After)
t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
  group_by(CFF.ID,Site,Before_After) %>%
  arrange(Site,CFF.ID,Before_After)
keep = most_recent %>% group_by(CFF.ID) %>% count() %>% filter(n == 2)
```

>>>>

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lmerTest)
library(nlme)
library(broom.mixed)
library(segmented)
library(performance)
library(AGD)
library(rspiro)
library(lubridate)
library(plotly)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
source("~/GitHub/CF/get_genotype_class_severity.R")
knitr::opts_knit$set(root.dir = home_dir)
```

@@

df$BMI.Percentile = pnorm(df$BMIz)*100
df$Weightz = y2z(df$Weight,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.wgt)

@@

# Get genotype severity
df$Genotype1_sev = get_cf_genotype_class_severity(df$Genotypes1)
df$Genotype2_sev = get_cf_genotype_class_severity(df$Genotypes2)

@@

# T1D duration
df$CFRD.Diagnosis.Date = mdy(df$CFRD.Diagnosis.Date)
df$T1D.Duration = as.numeric(difftime(df$Date,df$CFRD.Diagnosis.Date,units = "days"))/365.25

@@

```{r}
# Most recent before/after - limit to 2 years before start
# Want the value closest to start date before start, and most recent with respect to now from start
# (i.e. the longest time after start)
most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
  filter(Time > -2*365) %>%
  mutate(time_new = max(Time) - Time) %>%
  arrange(time_new) %>% 
  summarise(Age = first(na.omit(Age)),BMI.Percentile = first(na.omit(BMI.Percentile)),
            Weight = first(na.omit(Weight)),Sex = first(na.omit(Sex)),
            T1D.Duration = first(na.omit(T1D.Duration)),
            G.tube.in.past.12.months = first(na.omit(G.tube.in.past.12.months)),
            Pancreatic.Status = first(na.omit(Pancreatic.Status)),
            Genotype1_sev = first(na.omit(Genotype1_sev)),
            Genotype2_sev = first(na.omit(Genotype2_sev)),
            BMIz = first(na.omit(BMIz)),Weightz = first(na.omit(Weightz)),
            FEV1 = first(na.omit(FEV1)),FVC = first(na.omit(FVC)),
            HbA1c = first(na.omit(HbA1c)),.groups = "keep") %>%
  arrange(Site,CFF.ID,Before_After)
# # Check
# t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
#   group_by(CFF.ID,Site,Before_After) %>%
#   arrange(Site,CFF.ID,Before_After)
# Filter those without CFRD, and those without pre AND post
keep = most_recent %>% filter(!is.na(HbA1c)) %>%
  group_by(CFF.ID) %>% summarise(n = n()) %>%
  filter(n == 2) %>% pull(CFF.ID)
cfrd = unique(df$CFF.ID[!is.na(df$CFRD.Status) | !is.na(df$CFRD.Diagnosis.Date)])
keep = unique(keep,cfrd)
most_recent = most_recent %>% filter(CFF.ID %in% keep)
```
